---
output: html_document
---

# `r html.title`

## Introduction

**Human Activity Recognition** is a research field that is gaining attention.
This is due to emergence of technology that makes it possible to easily collect
these kinds of data using *activity trackers*.
  
The aim of this analysis is to predict how well users have executed a
particular exercise. The data used in this document is the
**Weight Lifting Exercises Dataset** from the [**Human Activity Recognition**
project](http://groupware.les.inf.puc-rio.br/har).

## Weight Lifting Exercises Dataset

For the purposes of this program, the data is split into two types: the
training and the test set. Both data are stored into one **csv** each.

## predict.exercise.R

All of the functions described in this document are implemented in the
[**predict.exercise.R**](predict.exercise.R) file located in this git
repository.

### Loading the Data

The **csv**'s are first loaded using the *load.data()* function. This function
calls *read.csv()* and sets the **na.string** so it would treat empty strings
as NA.

The types of data stored in both sets are similar in most cases, other than the
training set containing the **classe** variable, and the test set containing
the **problem_id**.

The **classe** variable pertains to the manner in which the exercise has been
performed. It contains five value. **Class A** corresponds to the exercise done
correctly, while the other values correspond to common errors of the exercise.
It is the goal of this analysis to predict this variable.

### Preprocessing the Data

The dataset also contains metadata on the data gathered. The metadata includes
the names of the participants, as well as the time and date they were gathered.
For the purposes of this analysis, we will exclude these variables from the
prediction.

```{r excluded data}
excluded.columns
```

Some data are also a summary of the data gathered for that window. These data
are discarded for the purposes of this analysis. The data will be NA for most
data other than those that are of a new window. Preprocessing is done by the
*preprocess.data()* function.

### Creating the Model

The model is created by the *create.model()* function.

```{r show create model}
create.model
```

The function first loads the training data, then preprocesses the data so only
the relevant data will be left. It will then use the *randomForest()* algorithm
to create the prediction model, modelling **classe** against all the other
variables.

### Predicting on the Test Data

The predictions on the test data is done by the *predict.data()* function.

```{r show predict data}
predict.data
```

The function loads and preprocesses the test data in the same manner as the
training data. It will then use the prediciton model passed to it to predict
the **classe** of the test data.

## Results

### Prediction Model

The generated prediction model had a
`r (1 - prediction.model$err.rate[nrow(prediction.model$err.rate), 1]) * 100`%
out-of-bag success rate. The resulting confusion matrix is shown below.

```{r model}
prediction.model$confusion
```

### Predictions on the Test Data

The resulting predictions are shown below.

```{r prediction}
prediction
```

These predictions are correct, as verified by the course's [submissions page]
(https://class.coursera.org/predmachlearn-010/assignment).
